---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
unzip("activity.zip")
activity <- read.csv("./activity.csv", colClasses = c("integer","Date","numeric"), na.strings = "NA")

```

## What is mean total number of steps taken per day?
Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
totalsteps <- aggregate(steps ~ date, data = activity,  sum, na.rm = T)
hist(totalsteps$steps, xlab = "Total steps per day", main = "Histogram of totalsteps/day")

```
Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
meansteps <- mean(totalsteps$steps)
mediansteps <- median(totalsteps$steps)
meansteps
mediansteps
```

## What is the average daily activity pattern?
Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE}
avestep_perinterval <- aggregate(steps~interval, data=activity, mean, na.rm=TRUE)
with(avestep_perinterval, plot(interval, steps, type = "l", main = "Average steps for each interval"))

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r, echo=TRUE}
sortsteps <- avestep_perinterval[order(-avestep_perinterval$steps),]
top5intervals <- subset(sortsteps[1:5,])$interval
top5intervals

```

## Imputing missing values
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo=TRUE}
totalmiss <- sum(is.na(activity))
totalmiss

```

Devise a strategy for filling in all of the missing values in the dataset (use the mean for that 5-minute interval). Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo=TRUE}
imputed <- activity 
for (i in 1:nrow(imputed)) {
  if (is.na(imputed$steps[i])) {
    imputed$steps[i] <- avestep_perinterval[which(imputed$interval[i] == avestep_perinterval$interval), ]$steps
  }
}
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
imputed_totalsteps <- aggregate(steps ~ date, data = imputed,  sum, na.rm = T)
hist(imputed_totalsteps$steps, xlab = "Total steps per day", main = "Histogram of totalsteps/day")

meansteps <- mean(imputed_totalsteps$steps)
mediansteps <- median(imputed_totalsteps$steps)
meansteps
mediansteps
```

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r, echo=TRUE}
imputed$day <- weekdays(imputed$date)
for (i in 1:nrow(imputed)) {
  if (imputed$day[i] == "Saturday" | imputed$day[i] == "Sunday") {
    imputed$day[i] = "Weekend"
  } else {
    imputed$day[i] = "Weekday"
  }
}

```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r, echo=TRUE}
library(lattice)
avestep_perinterval <- aggregate(steps~interval + day, data=imputed, mean)
xyplot(avestep_perinterval$steps ~ avestep_perinterval$interval | avestep_perinterval$day, layout=c(1,2), type = "l", xlab = "Interval", ylab = "Number of steps")

```

